---
title: "traitdata: Easy access to various ecological trait data"
output: github_document
editor_options: 
  chunk_output_type: console
---

## Installation

You can install traitdata from Github with:

```{r gh-installation, eval=F}
# Install devtools if not available
if(!"remotes" %in% installed.packages()[,"Package"]) install.packages("remotes")

# Install traitdata package from Github
remotes::install_github("RS-eco/traitdata", build_vignettes = T, force=T)
```

After installation, simply load the `traitdata` package:

```{r, eval=F}
library(traitdata)
```

**If you encounter a bug or if you have any problems, please file an [issue](https://github.com/RS-eco/traitdata/issues) on Github.**

## Data overview

There are 32 different data sets, which are included in this package:

datasetName	| basisOfRecord	| rightsHolder	     | DOI
----------- | ------------- | ------------------ | -----------
amniota | traitDatabase | Myrhvold et al. 2016 | [10.1890/15-0846R.1](http://doi.org/10.1890/15-0846R.1)
amphi_lifehist | traitDatabase | Trochet et al. 2014 | [10.3897/BDJ.2.e4123](http://doi.org/10.3897/BDJ.2.e4123)
amphibio | traitDatabase | Oliveira et al. 2017 | [10.1038/sdata.2017.123](http://doi.org/10.1038/sdata.2017.123)
an_age | traitDatabase | Tacutu et al. 2018 | [10.1093/nar/gkx1042](http://doi.org/10.1093/nar/gkx1042)
anuran_morpho | traitDatabase | Mendoza-Henao et al. 2019 | [10.1002/ecy.2685](http://doi.org/10.1002/ecy.2685)
arthropods | traitDatabase | Gossner et al. 2015 | [10.1038/sdata.2015.13](http://doi.org/10.1038/sdata.2015.13)
atlantic_birds | traitDatabase | Rodrigues et al. 2019 | [10.1002/ecy.2647](https://doi.org/10.1002/ecy.2647)
australian_birds | traitDatabase | Garnett et al. 2015 | [10.1038/sdata.2015.61](https://doi.org/10.1038/sdata.2015.61)
AvianBodySize | traitDatabase | Lislevand et al. 2007 | [10.1890/06-2054](https://doi.org/10.1890/06-2054)
AVONET | traitDatabase | Tobias et al. 2021 | 
[10.6084/m9.figshare.16586228.v2](https://doi.org/10.6084/m9.figshare.16586228.v2)
bird_behav | traitDatabase | Tobias & Pigot 2019 | [10.1098/rstb.2019.0012](http://dx.doi.org/10.1098/rstb.2019.0012)
carabids | traitDatabase | van der Plas et al. 2017 | [10.5061/dryad.53ds2](http://doi.org/10.5061/dryad.53ds2)
climber | traitDatabase | Schweiger et al. 2014 | [10.3897/zookeys.367.6185](http://doi.org/10.3897/zookeys.367.6185)
disperse | traitDatabase | Sarremejane et al. 2020 | [10.1038/s41597-020-00732-7](http://doi.org/10.1038/s41597-020-00732-7)
elton_birds	| traitDatabase	| Wilman et al. 2014 | [10.1890/13-1917.1](http://doi.org/10.1890/13-1917.1)
elton_mammals |	 traitDatabase	| Wilman et al. 2014 | [10.1890/13-1917.1](http://doi.org/10.1890/13-1917.1)
epiphytes | traitDatabase | Hietz et al. 2021 | [10.1111/1365-2745.13802](https://doi.org/10.1111/1365-2745.13802)
eubirds | traitDatabase | Storchová & Hořák 2017 | [10.1111/geb.12709](https://doi.org/10.1111/geb.12709)
fishmorph | traitDatabase | Brosse et al. 2021 | [10.1111/geb.13395](https://doi.org/10.1111/geb.13395)
globalHWI | traitDatabase | Sheard et al. 2020 | [10.1038/s41467-020-16313-6](http://doi.org/10.1038/s41467-020-16313-6)
globTherm | traitDatabase | Bennett et al. 2018 | [10.1038/sdata.2018.22](https://doi.org/10.1038/sdata.2018.22)
heteroptera | traitDatabase | Gossner et al. 2016 | [10.6084/m9.figshare.c.3307611.v1](https://doi.org/10.6084/m9.figshare.c.3307611.v1)
heteropteraRaw | traitDatabase | Gossner et al. 2016 | [10.6084/m9.figshare.c.3307611.v1](https://doi.org/10.6084/m9.figshare.c.3307611.v1)
lizard_traits | traitDatabase | Meiri 2018 | [10.1111/geb.12773](http://doi.org/10.1111/geb.12773)
mammal_diet | traitDatabase | Kissling et al. 2014 | [10.1002/ece3.1136](http://doi.org/10.1002/ece3.1136)
mammal_diet2 | traitDatabase | Gainsbury et al. 2018 | [10.1111/mam.12119](http://doi.org/10.1111/mam.12119)
marsupials | traitDatabase | Fisher et al. 2001 | [10.1890/0012-9658(2001)082[3531:TEBOLH]2.0.CO;2](https://doi.org/10.1890/0012-9658(2001)082[3531:TEBOLH]2.0.CO;2)
migbehav_birds	| literatureData	| Eyres & Fritz	| [10.12761/SGN.2017.10058](http://doi.org/10.12761/SGN.2017.10058)
pantheria | traitDatabase | Jones et al. 2009 | [10.1890/08-1494.1](http://doi.org/10.1890/08-1494.1)
passerines | traitDatabase | Ricklefs 2017 | [10.1002/ecy.1783](http://doi.org/10.1002/ecy.1783)
primates | traitDatabase | Galán-Acedo et al. 2020 | [10.1038/s41597-019-0059-9](http://doi.org/10.1038/s41597-019-0059-9)
reptile_lifehist | traitDatabase | Grimm et al. 2014 | [10.3897/natureconservation.9.8908](http://doi.org/10.3897/natureconservation.9.8908)
tetra_density | traitDatabase | Santini et al. 2018 | [10.1111/geb.12756](https://doi.org/10.1111/geb.12756)

**Note:** The code for how these datatsets were downloaded and processed can be found in the [data-raw](https://github.com/RS-eco/traitdata/tree/main/data-raw) folder.

**See also https://opentraits.org/datasets.html for an extensive list of Trait datasets.**

An overview of the different datasets can also be found here: `vignette("data_info")`. 

## Variables

* All published datasets contain their original variables and species names. 
Species names were split into 2 columns (Genus and Species) and where applicable into a 3rd column with Subspecies names.
* All datasets also have a scientificNameStd column, which is a standardised scientific name, 
in order to be able to merge the different datasets by species.

**Note:** Not all species names could be standardised, therefore some data entries might not contain a scientificNameStd value, please then refer to the Genus and Species column. In case you are interested in which species could not be standardised have a look at the `names_nonStd` file. Please also check out the `synonyms` data file, for species where an alternative name has been used for standardising the scientific name.

An overview of all variables with a description of each variable can be found in the `trait_glossary` data file:

```{r}
data(trait_glossary)
```

or in the `glossary` vignette:

```{r, eval=F}
vignette("trait_glossary")
```

## Data query

To connect to one or more datasets, we simply use the `data()` function.

```{r}
# Load Elton Traits
data("elton_birds")
```

Now we can use standard R calls to have a look at the data. 

First, we look at the column names of our dataset.

```{r}
# Look at the variable names
colnames(elton_birds)
```

Then, we check the class and first 6 rows of the first 10 columns of the `elton_traits` dataset:

```{r}
class(elton_birds)
head(elton_birds[,1:10])
```

For more information on how to use the data within the package, check out the `access-data` and all the other vignettes.

```{r, eval=F}
vignette("access-data")
```

Additional examples of how to use the different trait datasets can be found in the following vignettes:

```{r, eval=F}
vignette("island-birds")
vignette("migbehav_birds")
vignette("morpho-indices")
vignette("pantheria")
vignette("passerines")
```
