---
title: "Morphological indices"
author: "RS-eco"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Morphological indices}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Morphological indices

Morphological indices can be helpful to depict ecological differences between feeding guilds and between species differing in their preferred foraging stratum. 

There are 6 morphological indices, which are often calculated:

* Kipp’s index (Kipp’s distance divided by wing length), 
* tail-wing index (the longest rectrix divided by wing length), 
* bill height index (bill height divided by bill length), 
* bill width index (bill width divided by bill length), 
* lateral tarsus index (lateral tarsus diameter divided by tarsus length) and 
* sagittal tarsus index (sagittal tarsus diameter divided by tarsus length).

Here, we show how to calculate these indices, using the data included in this package:

```{r}
#' First we load our Data
library(traitdata)
data(passerines)

#' Secondly, we need to load the dplyr library
library(dplyr)

#' Now, we can calculate the indices using the mutate function from the dplyr package
#mutate(KippData, kipp_index = kipp / Wing)
#mutate(KippData, tail-wing_index = tail / Wing)
#mutate(KippData, bill_height_index = billheight/ billlength)
#mutate(KippData, bill_width_index = billwidth / billlength)
#mutate(KippData, lateral_tarsus_index = lateral.tarsus.diameter / tarsus.length)
#mutate(KippData, sagittal_tarsus_index = sagittal.tarsus.diamter / tarsus.length)
```

Now, we want to plot these indices:

```{r}
#' Load the ggplot2 library
library(ggplot2)

#' Plot a histogram of the 6 indices

```

